---
title: "A1: initial tutorial for biocBiocypher"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{A1: initial tutorialfor biocBiocypher}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

```{r setup,echo=FALSE,message=FALSE,results="hide"}
library(biocBiocypher)
library(BiocStyle)
library(DT)
```

# Introduction

This vignette goes through material at the "Adding data"
[tutorial document](https://biocypher.org/tutorial.html#section-1-adding-data).
Our objective is to use R and python together, with `r Biocpkg("basilisk")`
managing the python infrastructure.

We will simulate records with schematized information about 10 proteins and then
use `write_nodes` to generate a CSV file.  

# The primary interface

`loadBiocypher` connects the Biocypher modules to R via basilisk and
reticulate.  A completely isolated miniconda environment, currently using
Python 3.9, manages all the python code.

```{r dobas}
library(biocBiocypher)
bcobj = loadBiocypher()
bcobj
```

## The data generator

```{r lkgen}
gen = bcobj$generator_ref
names(gen)
```

The following R code generates records on 10 proteins:
```{r dolapp}
prots = lapply(1:10, function(x) gen$Protein())
names(prots[[1]])
prots[[1]]$properties
```

This list is not known to the python main module (`__main__`) however.  We need to use

```{r dopy}
reticulate::py_run_string("proteins = [Protein() for _ in range(10)]")
names(reticulate::py)  # symbols known to main
```

## Producing the graph nodes

Several configuration files are defined for this specific tutorial.
```{r config}
bc_config_path = system.file("tutorial_0.5.11", 
     "01_biocypher_config.yaml", package="biocBiocypher")
schema_config_path = system.file("tutorial_0.5.11", 
     "01_schema_config.yaml", package="biocBiocypher")
readLines(schema_config_path)
```

These configurations are loaded into the main
interface.  We "update" the YAML in `bc_config_path`
so that the output folder is user-selectable.
The default output folder is a temporary folder.

```{r tomain}
bc = bcobj$biocypher_ref
bc_configd = bc$BioCypher(
    biocypher_config_path=update_bc_config(bc_config_path),
    schema_config_path=schema_config_path
)
```
The `node_generator` was written to use a globally defined variable `proteins`.
That was defined above with `py_run_string`.

We write out the nodes:

```{r look}
bc_configd$write_nodes(gen$node_generator())
```

We can retrieve the configured output directory from `bc_configd`.
In this case the files are a 'header' and a semicolon-delimited
data file.  We parse and put them together in the following,
then create a searchable HTML table.
```{r dod}
od= bc_configd$base_config$output_directory
fi = dir(od, full=TRUE,patt="part")
he = strsplit(readLines(dir(od, full=TRUE, patt="head"), warn=FALSE), ";")[[1]]
dat = read.delim(fi, sep=";", h=FALSE)
names(dat) = he
library(DT)
datatable(dat)
```



```{r domo}
cat(reticulate::py_capture_output(bc_configd$summary()))
```

# Summary

The main Biocypher processes we have examined thus far include

- configuring metadata about protein annotations to be handled, in YAML

```
protein:                            # mapping
  represented_as: node              # schema configuration
  preferred_id: uniprot             # uniqueness
  input_label: uniprot_protein      # connection to input stream
```

- configuring the "back end", in this case 'offline', also in YAML


- transforming a stream of protein identifier data into graph nodes
```
bc_configd$write_nodes(gen$node_generator())
```

The node information was serialized to a tabular format.

In the next tutorial we will combine information from different
annotation types.


# Session information

```{r lksess}
sessionInfo()
```




