---
title: "A2: using biocBiocypher to unite information from different namespaces"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{A2: using biocBiocypher to unite information from different namespaces}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

```{r setup,echo=FALSE,message=FALSE,results="hide"}
library(biocBiocypher)
library(BiocStyle)
library(DT)
```

# Introduction

This vignette moves beyond the A1 vignette, which simply imported
information on simulated proteins identified by Uniprot identifiers.

Now we will use two identifier types: Uniprot and NCBI Entrez.

# Setup


```{r dobas}
library(biocBiocypher)
bcobj = loadBiocypher()
bcobj
```

## The data generator

```{r lkgen}
gen = bcobj$generator_ref
names(gen)
```

In the previous vignette, we used `Protein()`.
Now we will use `EntrezProtein()`.

```{r dolapp}
prots = lapply(1:10, function(x) gen$EntrezProtein())
names(prots[[1]])
prots[[1]]$properties
```

The python code to work with Entrez and Uniprot
together is:
```{r litco}
twop = "proteins = [
    p for sublist in zip(
        [Protein() for _ in range(10)],
        [EntrezProtein() for _ in range(10)],
    ) for p in sublist
]"
tt = tempfile()
writeLines(twop, tt)
```

We run the code to get the protein metadata and sequences.

```{r dopy}
reticulate::py_run_file(tt)

names(reticulate::py)  # symbols known to main
```

## Producing the graph nodes

As before, we use
the configuration files defined for this specific tutorial.
```{r config}
bc_config_path = system.file("tutorial_0.5.11", 
     "02_biocypher_config.yaml", package="biocBiocypher")
schema_config_path = system.file("tutorial_0.5.11", 
     "02_schema_config.yaml", package="biocBiocypher")
readLines(schema_config_path)
```

These configurations are loaded into the main
interface.  We "update" the YAML in `bc_config_path`
so that the output folder is user-selectable.
The default output folder is a temporary folder.

```{r tomain}
bc = bcobj$biocypher_ref
bc_configd = bc$BioCypher(
    biocypher_config_path=update_bc_config(bc_config_path),
    schema_config_path=schema_config_path
)
```
The `node_generator` was written to use a globally defined variable `proteins`.
That was defined above with `py_run_string`.

We write out the nodes:

```{r look}
bc_configd$write_nodes(gen$node_generator())
```

We retrieve the configured output directory from `bc_configd`.
The files are again a 'header' and a semicolon-delimited
data file.  We parse and put them together in the following,
then create a searchable HTML table.
```{r dod}
od= bc_configd$base_config$output_directory
fi = dir(od, full=TRUE,patt="part")
he = strsplit(readLines(dir(od, full=TRUE, patt="head"), warn=FALSE), ";")[[1]]
dat = read.delim(fi, sep=";", h=FALSE)
names(dat) = he
library(DT)
datatable(dat)
```



```{r domo}
cat(reticulate::py_capture_output(bc_configd$summary()))
```

# Summary

Data generation in this vignette uses two namespaces to produce
graph nodes.


# Session information

```{r lksess}
sessionInfo()
```




